---
title: "Analysis report"
subtitle: 'Supplement to "Efficacy of surgery and bronchoscopic resection in adult patients with resectable pulmonary typical carcinoid tumours"'
author: "Andrés Obeso"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
library(databrew)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

library(knitr)
# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = F,
               fig.width = 7,
               fig.height = 6)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})

```

# Summary

This report summarizes (in both code and text) the analysis carried out for "Efficacy of surgery and bronchoscopic resection in adult patients with resectable pulmonary typical carcinoid tumours: a systematic review." All code is available online.^[http://www.github.com/joebrew/andresobeso]


# Data processing

Prior to analysis, data from the excel spreadsheet (`data/Tabla excel analisis.xlsx`) are read in and processed to a "tidy" format [@tidy] using the script `global.R`.

```{r}
source('global.R')
```

# Dava visualization

## Periop complications over time

```{r}
ggplot(data = df,
       aes(x = `Follow up (months)`,
           y = `Periop complications (%)`,
           color = Group)) +
  geom_point(aes(size = N)) + 
  theme_databrew()
```

## Postoperative mortality (0%) over time

```{r}
ggplot(data = df,
       aes(x = `Follow up (months)`,
           y = `Postoperative mortality (0%)`,
           color = Group)) +
  geom_point(aes(size = N)) + 
  theme_databrew()
```

## Complete resection (%) over time

```{r}
ggplot(data = df,
       aes(x = `Follow up (months)`,
           y = `Complete resection (%)`,
           color = Group)) +
  geom_point(aes(size = N)) + 
  theme_databrew()
```

## Residual disease (%) over time

```{r}
ggplot(data = df,
       aes(x = `Follow up (months)`,
           y = `Residual disease (%)`,
           color = Group)) +
  geom_point(aes(size = N)) + 
  theme_databrew()
```

## Recurrence (%) over time

```{r}
ggplot(data = df,
       aes(x = `Follow up (months)`,
           y = `Recurrence (%)`,
           color = Group)) +
  geom_point(aes(size = N)) + 
  geom_smooth() +
  theme_databrew()
```

## Long term success rate

```{r}
ggplot(data = df,
       aes(x = `Group`,
           y = `Long term success rate`,
           color = Group)) +
  geom_violin(alpha = 0.8) +
  geom_jitter(aes(size = N), alpha = 0.8) + 
  theme_databrew()
```

## 5 year survival rate

```{r}
ggplot(data = df,
       aes(x = `Group`,
           y = `5 year survival`,
           color = Group)) +
  geom_violin(alpha = 0.8) +
  geom_jitter(aes(size = N), alpha = 0.8) + 
  theme_databrew()
```

## 10 year survival rate

```{r}
ggplot(data = df,
       aes(x = `Group`,
           y = `10 year survival`,
           color = Group)) +
  geom_violin(alpha = 0.8) +
  geom_jitter(aes(size = N), alpha = 0.8) + 
  theme_databrew()
```

## Long term success rate by sex

```{r}
ggplot(data = df,
       aes(x = `Male (%)`,
           y = `Long term success rate`)) +
  geom_smooth(alpha = 0.8) + 
  geom_point(alpha = 0.8) +
  theme_databrew()
```


## Long term success rate by age

```{r}
ggplot(data = df,
       aes(x = `Age (yr)`,
           y = `Long term success rate`)) +
  geom_smooth(alpha = 0.8) + 
  geom_point(alpha = 0.8) +
  theme_databrew()
```


# Models

Are there statistically significant differences in the following outcome measures, as a function of the intervention type:
Complete Resection, Residual Disease, Recurrence rate, Long tern successful rate and Survival

## Methods for modeling

Simple ordinary least squares, weighted by `N` (sample size), and adjusted for Age, Sex, and months followed up. Formulaically:

```
<outcome> ~ (Group + `Age (yr)` + `Male (%)` + `Follow up (months)`) * `N`
```

The results are both tabular and graphical. In both, the "reference" group is Surgery (ie, the value shown is for Bronchoscopy). A p-value of less than 0.05, or a chart with the vertical uncertainty line not crossing the 0 line, is considered statistically significant. For survival outcomes, follow up months was removed from the right hand side of the equation.

We use the below R code to generate a model and plot for each outcome variable.

```{r, echo = TRUE}
# Define function for model
model <- function(outcome,
                  data){
  data$Group <- factor(data$Group, levels = c('Surgery', 'Bronchoscopy'))
  if(grepl('urvival', outcome)){
    the_formula <- paste0('`', outcome, '` ~ Group +',  '`Age (yr)` + `Male (%)`')
  } else {
    the_formula <- paste0('`', outcome, '` ~ Group +',  '`Age (yr)` + `Male (%)` + `Follow up (months)`')
  }
  
  the_formula <- as.formula(the_formula)
  fit <- lm(the_formula, weights = data$N, data = data)
  ci <- confint(fit)
  coefs <- coef(fit)
  out <-data.frame(coef = names(coefs), coefs, ci)
  out$coef <- gsub('Group', '', out$coef)
  row.names(out) <- NULL
  names(out) <- c('Variable', 'Coefficient', 'Lower', 'Upper')
  out$`P Value` <- summary(fit)$coefficients[,4]
  out_list <- list()
  out_list[[1]] <- out
  g <- ggplot(data = out %>%
                filter(Variable != '(Intercept)'),
              aes(x = Variable,
                  y = Coefficient)) +
    geom_point(color = 'red', alpha = 0.7) +
    geom_linerange(aes(x = Variable,
                       ymax = Upper,
                       ymin = Lower),
                   alpha = 0.7) +
    geom_hline(yintercept = 0, alpha = 0.7, lty = 2) +
    theme_databrew() +
    labs(title = '')
  out_list[[2]] <- g
  names(out_list) <- c('data', 'plot')
  return(out_list)
}
```

## Complete Resection

```{r}
x <- model(outcome = 'Complete resection (%)',
           data = df)
```

```{r, results = 'asis'}
knitr::kable(x$data)
```

```{r}
x$plot
```


## Residual disease (%)

```{r}
x <- model(outcome = 'Residual disease (%)',
           data = df)
```

```{r, results = 'asis'}
knitr::kable(x$data)
```

```{r}
x$plot
```

## Recurrence (%)

```{r}
x <- model(outcome = 'Recurrence (%)',
           data = df)
```

```{r, results = 'asis'}
knitr::kable(x$data)
```

```{r}
x$plot
```


## Long term success rate

```{r}
x <- model(outcome = 'Long term success rate',
           data = df)
```

```{r, results = 'asis'}
knitr::kable(x$data)
```

```{r}
x$plot
```

## 5 year survival

```{r}
x <- model(outcome = '5 year survival',
           data = df)
```

```{r, results = 'asis'}
knitr::kable(x$data)
```

```{r}
x$plot
```

## 10 year survival

```{r}
x <- model(outcome = '10 year survival',
           data = df)
```

```{r, results = 'asis'}
knitr::kable(x$data)
```

```{r}
x$plot
```


```{r eval=FALSE}
file.edit(
  tufte:::template_resources(
    'tufte_html', '..', 'skeleton', 'skeleton.Rmd'
  )
)
```

```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
# knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```
